\thetitle

### 2008-06-15 paper.tex
Added basic behaviors table; sending to John.
Miscellaneous tweaking.

Our first example is the packets-bytes-duration (\caps{PBD}) representation.
The representation is $\pbd: \Traffic \to \R^3$, mapping
$\flow \mapsto \tuple{p,b,d} \in \R^3$, where
\begin{align}
p &= \pkts, &
b &= \sum_{i=1}^{\mathclap{p}}{\size_i}, &
d &= \sum_{i=1}^{\mathclap{p-1}}{\ival_i}.
\end{align}
This representation encapsulates the information necessary to reproduce a \caps{CBR} version of each flow:
packets have payloads of $\floor{\fracx{b}{p}}$ bytes, and are sent every $\fracx{d}{p}$ seconds.\pbdnote
Given two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, their sum as vectors, 
\begin{align}
\vec{x}_1+\vec{x}_2=\tuple{p_1+p_2,b_1+b_2,d_1+d_2},
\end{align}
provides an appropriate description of the aggregate behavior of the two flows together.
Similarly, their average, \latin{i.e.} $\frac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.
In contrast, a ``packets-payload-interval'' representation, expressing the \caps{CBR} parameters directly, does not have this property:
the sum of average payloads is not the overall average of payloads;
likewise for intervals.
In the \caps{PBD} representation, the linearity property leads to natural descriptions of aggregate behavior through vector addition, whereas in the packets-payload-interval representation, vector addition is ``unnatural.''

@@

Two other unspecified constants appear in \Table{quantization}: $\beta$ and $\gamma$.
These parameters are non-linear scaling exponents that allows the quantization to smoothly shift from fine-grained resolution at the low end of the spectrum to coarse-grained resolution at the high end.
We have chosen $\beta=\fracx{1}{2.7}$ and $\gamma=\fracx{1}{3}$ because these values yield desirable dynamic ranges: from single bytes at the lower end to megabytes at the high end for flow sizes; from microseconds to seconds for intervals.

@@

From the matrices $\mat{Z}$ and $\mat{V}$ we can compute a variety of statistics about packet sizes and inter-packet intervals.
The flows vectors in the matrices may represent individual flows or composites of many flows.
% about packet sizes, total packet count, inter-packet intervals, and total duration for each representation vector, which may represent a single actual flow, or the composite behavior of many flows.
% We begin by expressing the vectors for total bytes, total packets and total duration for a matrix of flow representations, with size and interval matrices $\mat{Z}$ and $\mat{V}$.
The computations require knowing the values that each size of interval matrix column represents.
% The computed duration is approximate because information is irretrievably lost in the quantization process.
The vectors of representative values for $\mat{Z}$ and $\mat{V}$ respectively are:
\begin{align}
\vec{\sigma} &= \tuple{1,2,3,\dots,d_z} \\
\vec{\tau}   &= \tuple{\Df_\ival\left(k-\fracx{1}{2}\right)}_{k=1}^{d_v}.
\end{align}
The packet count, total byte and total duration vectors are computed as
\begin{align}
\vec{\Sigma}_\pkts  &= \mat{Z}\trans{\ones} = \mat{V}\trans{\ones} \in \R^f \eqn{packet-count}\\
\vec{\Sigma}_\bytes &= \mat{Z}\trans{\vec{\sigma}} \in \R^f \eqn{total-bytes} \\
\vec{\Sigma}_\ivals &= \mat{V}\trans{\vec{\tau}} \in \R^f \eqn{total-duration}.
\end{align}
Here $\ones$ is a row vector of all ones.\onesnote
Packet count and total bytes values are exact;
the total durations are only approximate because information is irretrievably lost in the quantization process.
Using these vectors of totals, we may readily compute the average packet sizes and inter-packet intervals:
\begin{align}
\vec{\mean{z}} &=
  \vec{\Sigma}_\bytes \eldiv \vec{\Sigma}_\pkts
  \phantom{= \mat{V}\trans{\vec{\tau}} \eldiv \mat{V}\trans{\ones}}
  \mathllap{= \mat{Z}\trans{\vec{\sigma}} \eldiv \mat{Z}\trans{\ones}} \\
\vec{\mean{v}} &=
  \phantom{\vec{\Sigma}_\bytes \eldiv \vec{\Sigma}_\pkts}
  \mathllap{\vec{\Sigma_{\ivals}} \eldiv \vec{\Sigma_{\pkts}}}
  = \mat{V}\trans{\vec{\tau}} \eldiv \mat{V}\trans{\ones}.
\end{align}
The symbol $\eldiv$ here represents element-wise division of vectors, as it does in Matlab.

@@

Since uniform behaviors entail homogeneity of properties with respect to each flow individually, a trace matrix can be transformed to a matrix satisfying the model's assumptions via right matrix multiplication.
Right multiplication mixes the elements within each row, and if the mixing weights are uniform, the result is uniform behavior.
For example, to make the roles of the nodes with respect to each flow statistically identical, we transform $\mat{S}$ and $\mat{D}$:
\begin{align}
\mat{S}' &= \rStoch(\mat{S}\ones) = \smallfrac{1}{n} \mat{S}\ones \in \R^{f \times n} \\
\mat{D}' &= \rStoch(\mat{D}\ones) = \smallfrac{1}{n} \mat{D}\ones\in \R^{f \times n}.
\end{align}
The ``$\rStoch$'' operator scales each row to sum to unity, thus making the entire matrix row-stochastic;
in this case, this is equivalent to scaling by $\smallfrac{1}{n}$.
These transformations of $\mat{S}$ and $\mat{D}$ make all the values in each flow's row identically the average of the original $n$ values.
As another example, the constant bit-rate (\caps{CBR}) flow model, in its strictest sense, is an intra-flow model, which stipulates that all the packets in each flow behave identically, having identical sizes and intervals;
thus each flow's bit-rate is constant, as the name implies.
The strict \caps{CBR} model can be easily expressed with the packets-bytes-duration (\caps{PBD}) model used as an example in \Section{representations}.
The pair of matrices $[\mat{Z}~\mat{V}]$ is transformed to \caps{PBD} form using the vectors defined in \Section{flow-totals}:
$[\mat{Z}~\mat{V}]\,\mat{\Gamma}\in\R^{f \times 3}$, where
\begin{align}
\mat{\Gamma}=\begin{pmatrix}
\ones & \vec{\sigma} & \zeros \\
\zeros & \zeros & \vec{\tau}
\end{pmatrix} \in \R^{(d_v+d_z)\times{3}}.
\end{align}
\Figure{trace-vs-uniform} graphically depicts examples of uniform behaviors as compared to realistic behaviors from traces.

@@

Marginality is an orthogonal concept to uniformity:
the distribution of some property is identical \textit{across} all flows.
For example, if we determine a distribution of flow sizes across the entire network and assume that this distribution applies independent of other flow properties, this is a marginal model for flow size.
Marginal behavior is effected by left multiplication by a matrix with uniform columns.
This specific example is generated by left-uniformizing the flow size matrix:
$\mat{B}' = \smallfrac{1}{f}\ones\mat{B}$.
The traffic models proposed by {\FHC} are all marginal:
they propose certain network-wide parametric distributions for start times, flow sizes and the number of flows per node.
These models do not account for non-independence between these and other flow properties.
We will discuss this limitation and its implications when we analyze our experimental results.

@@

The striking results of applying nonnegative matrix factorization to $\mat{P}$ are shown in \Figure{basic-behaviors}. Twelve basic behaviors suffice to explain all the traffic in a ``toy''  sample of 4138 flows. The pairs of packet size and inter-packet interval distributions are almost immediately recognizable as resembling intuitive modes of network behavior. The first basic behavior, for example has all packets near the \caps{MTU}, while intervals are smoothly distributed around a fraction of a millisecond. This is what we would expect from downloading large files; and indeed, the breakdown of packets by port number bears out this interpretation---almost all of the traffic is \caps{HTTP} server traffic, with small slivers of \caps{IMAP} and \caps{SSH} server traffic. The second basic behavior is equally intuitive: tiny packets with frequency distributed around 0.1 seconds---this is all \caps{SSH} traffic, as indicated by the associate pie chart. This is precisely the behavior we expect from interacting with a remote terminal via keystrokes and screen updates.

% TODO: say something here...

% What are the ramifications of these results? In our previous work, we had already demonstrated the drastic distortions that are introduced by uniformity assumptions made by the most common traffic models. This work serves to further critique established common assumptions in traffic modeling. The strongest message to be taken away is that the aspects of traffic behavior of both nodes and flows \textit{must} be considered jointly. It is not sufficient to continue modeling marginal distributions of network properties. These marginal models do not lend them selves easily to traffic generation in the first place, and when marginality assumptions are applied to trace traffic, the result is a severe distortion of important network performance characteristics. It is now well established that network traffic patterns have an impact on network performance that cannot be ignored. Even experimental deployments cannot avoid the need for more realistic traffic workload models. While using a real, physical network successfully sidesteps simulation problems below the application layer, without realistic traffic models, reliable, meaningful performance predictions remain beyond our reach.

%\subsection{Generality of Results}

%% TODO: talk about other synthetic models not considered here.
%% The point of our analysis is not to show that no synthetic models can be accurate, but rather, that the commonly used models, and obvious variations thereof, are not accurate.

%The most significant limitation on the generality of this analysis is that it is based entirely on a single data set from \caps{IETF}60---albeit a large and varied one. It is possible that traffic in this trace happens to produce network performance that is unusually dissimilar to standard traffic models. This data set, however, represents a highly heterogeneous collection of network usage behaviors, from slow and steady off-peak usage, to extremely heavy peak usage: over 800 users, 33 thousand flows, and 1 million packets in a single 10-minute trace segment. Despite the broad variety of behaviors, the results are consistent: in all types of usage scenarios, simplistic traffic models, like uniform \caps{CBR}, systematically skew important performance measurements at all levels of the network. While the precise results for other data sets might differ, it is very unlikely that \caps{CBR} traffic models will happen to accurately reproduce realistic performance in other experiments. This paper provides strong evidence that better traffic models are needed for performance evaluations.

% \subsection{Towards Realistic Models of Wireless Workload}
% 
% What would better traffic models look like? How can we create them? One possible approach is to use actual traffic traces as we have done. This approach is unsatisfactory, however, because it provides the experimenter with almost no control over experiments. Synthetic models have parameters, which can be tweaked as necessary---adjusting, for example, the number of active nodes in a simulation, without affecting other parameters. Traffic traces, on the other hand, must be used without significant alteration if they are to actually provide the desired realism. This inability to control all but one or two parameters is nearly crippling for experimental purposes. Therefore, the research community needs models whose parameters can be adjusted independently, but whose performance characteristics can be made to resemble whatever example of network behavior one might need. The general matrix model does not yet fulfill this role since its parameters cannot be freely varied. We believe, however, that \caps{GMM} plays a vital role in the ongoing process of developing  of better models. By reducing the general problem of traffic modeling to the domain of linear algebra, we have provided traffic researchers with a vast array of powerful theoretical and computational tools that can be immediately applied to the problem at hand.\footnote{The derived models used in this paper were all implemented in Matlab, each in only a few simple lines. All the models used here are implemented in only 19 total lines of code. Of course, the framework to apply them to traffic traces and simulations is somewhat larger.} What are the next steps? The authors are currently investigating a variety of matrix factorization and clustering techniques to extract hidden structure from a large and diverse body of real traffic.

%The reader will note that the transformations applied in this paper are all quite simple. Our conditioning matrices primarily consisted of ones and zeros. The algebraic representation of complete workload

%The ``messiness'' of the performance comparison from trace data in Figure~\ref{fig:mac-control-overhead-cmp} illustrates why using traces directly is not ideal: each data point differs not only in the number of nodes shown on the $x$-axis, but also in other dimensions, such as the number of flows and packets, and the average flow duration. The result is a highly noisy comparison, affected by many unseen parameters. Only by applying a local smoothing algorithm are trends somewhat elucidated.

%Instead of using trace data directly, it should be possible to configure a synthetic traffic model based on observations from a real data set, and then run side-by-side simulations using the synthetic model and the real data, producing statistically equivalent performance results. This is precisely what our definition of sufficient realism entails. The work in this paper provides the tools to measure how close to this ideal a model is and in what areas it needs improvement. Without this feedback, any improvements in realism are purely guesswork. Our breakdown of traffic behavior into three orthogonal levels also allows the problem to be approached in smaller pieces, rather than being solved all at once.

%The next step towards better traffic models is to investigate which aspects of real traces may be altered without detrimentally affecting the resulting performance metrics. For example, to test whether a complex time-series model of packet behavior is necessary, we randomize the order of the packet sizes and/or inter-packet intervals and compare performance using these randomized traces against performance using the original traces. If the performance is unchanged, we can conclude that no complex time-series model of packet behavior is necessary: sampling the packet sizes and inter-packet intervals from empirical distributions is sufficiently realistic. If, on the other hand, the performance characteristics are altered by shuffling packets, then some time-series model of packet behavior is needed. By partially randomizing the packet order in specific ways, the exact limits of realism necessary can be found. A similar approach will allow the development of realistic models for the other levels of network usage behavior.

@@

%The real significance of this paper lies not in showing that the commonly used \caps{CBR} traffic model is unrealistic. Networking researchers understand that \caps{CBR} is at best a na\"ive approximation of real network traffic. Nor is the real significance that we have quantized the high degree of inaccuracy that can be induced by using this model. The real significance of this paper lies in providing an well-defined, objective measure of realism for traffic models. This has never previously existed: evaluations of realism have relied on essentially arbitrary statistical measures of similarity to real traffic. Existing traffic generators use empirical distributions of such quantities as file size, inter-connection time, packet size, or inter-packet intervals. They can only hope that reproducing these distributions as good enough. Without an external measure of accuracy, there's simply no way to know.

%attempting to showing that existing models are unrealistic, or that new models are more realistic have always focused on arbitrary statistical measures. A

@@

%This research rigorously quantifies the impact of a variety of synthetic traffic models on performance metrics that wireless researchers use to evaluate new technologies and protocols. The first step in this assessment process was to formally define what it means for a network usage model to be sufficiently realistic. In essence, a model is considered realistic if it produces performance results that are statistically equivalent to those produced by real usage.
%A well-defined, objective measure of realism for traffic models has not previously existed. Evaluations of realism have formerly relied on essentially arbitrary statistical measures of similarity to real traffic, which may or may not affect the performance metrics that researchers care about.
%The definition of sufficient realism leads us to our general experimental approach: we use differential analysis comparing performance metrics derived from real traffic with those derived from synthetic traffic models. The theoretical contributions of this analysis are:
%\begin{enumerate}
%\item An in-depth analysis of the desirable mathematical properties of a measure of error for performance metrics.
%\item Proof that the unique measure of error that satisfies these properties is the log-ratio of metric values.
%\item Three rigorous tests of statistical equivalence between synthetic and real performance results.
%\end{enumerate}
%These analytical tools allow the evaluation of realism over a collection of drastically different usage scenarios. Evaluation over a heterogeneous collection of scenarios is essential to establishing the credibility of usage models. Moreover, these theoretical results are equally applicable to other types of usage models---for example, mobility.

%On the practical side, this paper gives crucial insight into why most researchers do not trust simulation results: with the traffic models commonly used, the results are unlikely to reflect real performance. Moreover, this problem will also hamper experiments in test-bed wireless networks, so long as the same na\"ive workload models are used. The only way to address this fundamental lack of realism is to develop usage models that reproduce important performance metrics more accurately. Our theoretical results provide the tools necessary to do this. The development of better traffic models should begin with real traces, and proceed by incremental changes, checked by differential analysis.
%%First, alter a small aspect of the trace, simulate, then compare. If the realism of the results is unaffected, the traffic feature altered was inessential. Otherwise, it is a feature of behavior that must be captured in a realistic traffic model.
%This approach will allow the precise mapping of which aspects of traffic patterns have an impact on performance and which ones can be safely abstracted away.
%% TODO: maybe cut out some of this last bit.

@@

% Hernandez06:dissertation: ``complete methodology for reproducing the traffic observed on a network link in a closed-loop manner, and proposed a number of metrics for studying the realism of the generated traffic.''

\vfill
\pagebreak
\bibliographystyle{plain}
\bibliography{references}

>>>>

Our first example is the packets-bytes-duration (\caps{PBD}) representation.
The representation is $\pbd: \Traffic \to \R^3$, mapping
$\flow \mapsto \tuple{p,b,d} \in \R^3$, where
\begin{align}
p &= \pkts, &
b &= \sum_{i=1}^{\mathclap{p}}{\size_i}, &
d &= \sum_{i=1}^{\mathclap{p-1}}{\ival_i}.
\end{align}
This representation encapsulates the information necessary to reproduce a \caps{CBR} version of each flow:
packets have payloads of $\floor{\fracx{b}{p}}$ bytes, and are sent every $\fracx{d}{p}$ seconds.\pbdnote
Given two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, their sum as vectors, 
\begin{align}
\vec{x}_1+\vec{x}_2=\tuple{p_1+p_2,b_1+b_2,d_1+d_2},
\end{align}
provides an appropriate description of the aggregate behavior of the two flows together.
Similarly, their average, \latin{i.e.} $\smallfrac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.
In contrast, a ``packets-payload-interval'' representation, expressing the \caps{CBR} parameters directly, does not have this property:
the sum of average payloads is not the overall average of payloads;
likewise for intervals.
In the \caps{PBD} representation, the linearity property leads to natural descriptions of aggregate behavior through vector addition, whereas in the packets-payload-interval representation, vector addition is ``unnatural.''

@@

Two other unspecified constants appear in \Table{quantization}: $\beta$ and $\gamma$.
These parameters are non-linear scaling exponents that allows the quantization to smoothly shift from fine-grained resolution at the low end of the spectrum to coarse-grained resolution at the high end.
We have chosen $\beta=\nicefrac{1}{2.7}$ and $\gamma=\nicefrac{1}{3}$ because these values yield desirable dynamic ranges: from single bytes at the lower end to megabytes at the high end for flow sizes; from microseconds to seconds for intervals.

@@

From the matrices $\mat{Z}$ and $\mat{V}$ we can compute a variety of statistics about packet sizes and inter-packet intervals.
The flows vectors in the matrices may represent individual flows or composites of many flows.
% about packet sizes, total packet count, inter-packet intervals, and total duration for each representation vector, which may represent a single actual flow, or the composite behavior of many flows.
% We begin by expressing the vectors for total bytes, total packets and total duration for a matrix of flow representations, with size and interval matrices $\mat{Z}$ and $\mat{V}$.
The computations require knowing the values that each size of interval matrix column represents.
% The computed duration is approximate because information is irretrievably lost in the quantization process.
The vectors of representative values for $\mat{Z}$ and $\mat{V}$ respectively are:
\begin{align}
\vec{\sigma} &= \tuple{1,2,3,\dots,d_z} \\
\vec{\tau}   &= \tuple{\Df_\ival\left(k-\smallfrac{1}{2}\right)}_{k=1}^{d_v}.
\end{align}
The packet count, total byte and total duration vectors are computed as
\begin{align}
\vec{\Sigma}_\pkts  &= \mat{Z}\trans{\ones} = \mat{V}\trans{\ones} \in \R^f \eqn{packet-count}\\
\vec{\Sigma}_\bytes &= \mat{Z}\trans{\vec{\sigma}} \in \R^f \eqn{total-bytes} \\
\vec{\Sigma}_\ivals &= \mat{V}\trans{\vec{\tau}} \in \R^f \eqn{total-duration}.
\end{align}
Here $\ones$ is a row vector of all ones.\onesnote
Packet count and total bytes values are exact;
the total durations are only approximate because information is irretrievably lost in the quantization process.
Using these vectors of totals, we may readily compute the average packet sizes and inter-packet intervals:
\begin{align}
\vec{\mean{z}} &=
  \vec{\Sigma}_\bytes \eldiv \vec{\Sigma}_\pkts
  \phantom{= \mat{V}\trans{\vec{\tau}} \eldiv \mat{V}\trans{\ones}}
  \mathllap{= \mat{Z}\trans{\vec{\sigma}} \eldiv \mat{Z}\trans{\ones}} \\
\vec{\mean{v}} &=
  \phantom{\vec{\Sigma}_\bytes \eldiv \vec{\Sigma}_\pkts}
  \mathllap{\vec{\Sigma_{\ivals}} \eldiv \vec{\Sigma_{\pkts}}}
  = \mat{V}\trans{\vec{\tau}} \eldiv \mat{V}\trans{\ones}.
\end{align}
The symbol $\eldiv$ here represents element-wise division of vectors, as it does in Matlab.

@@

Since uniform behaviors entail homogeneity of properties with respect to each flow individually, a trace matrix can be transformed to a matrix satisfying the model's assumptions via right matrix multiplication.
Right multiplication mixes the elements within each row, and if the mixing weights are uniform, the result is uniform behavior.
For example, to make the roles of the nodes with respect to each flow statistically identical, we transform $\mat{S}$ and $\mat{D}$:
\begin{align}
\mat{S}' &= \rStoch(\mat{S}\ones) = \smallfrac{1}{n} \mat{S}\ones \in \R^{f \times n} \\
\mat{D}' &= \rStoch(\mat{D}\ones) = \smallfrac{1}{n} \mat{D}\ones\in \R^{f \times n}.
\end{align}
The ``$\rStoch$'' operator scales each row to sum to unity, thus making the entire matrix row-stochastic;
in this case, this is equivalent to scaling by $\nicefrac{1}{n}$.
These transformations of $\mat{S}$ and $\mat{D}$ make all the values in each flow's row identically the average of the original $n$ values.
As another example, the constant bit-rate (\caps{CBR}) flow model, in its strictest sense, is an intra-flow model, which stipulates that all the packets in each flow behave identically, having identical sizes and intervals;
thus each flow's bit-rate is constant, as the name implies.
The strict \caps{CBR} model can be easily expressed with the packets-bytes-duration (\caps{PBD}) model used as an example in \Section{representations}.
The pair of matrices $[\mat{Z}~\mat{V}]$ is transformed to \caps{PBD} form using the vectors defined in \Section{flow-totals}:
$[\mat{Z}~\mat{V}]\,\mat{\Gamma}\in\R^{f \times 3}$, where
\begin{align}
\mat{\Gamma}=\begin{pmatrix}
\ones & \vec{\sigma} & \zeros \\
\zeros & \zeros & \vec{\tau}
\end{pmatrix} \in \R^{(d_v+d_z)\times{3}}.
\end{align}
\Figure{trace-vs-uniform} graphically depicts examples of uniform behaviors as compared to realistic behaviors from traces.

@@

Marginality is an orthogonal concept to uniformity:
the distribution of some property is identical \textit{across} all flows.
For example, if we determine a distribution of flow sizes across the entire network and assume that this distribution applies independent of other flow properties, this is a marginal model for flow size.
Marginal behavior is effected by left multiplication by a matrix with uniform columns.
This specific example is generated by left-uniformizing the flow size matrix:
$\mat{B}' = (\nicefrac{1}{f})\,\ones\mat{B}$.
The traffic models proposed by {\FHC} are all marginal:
they propose certain network-wide parametric distributions for start times, flow sizes and the number of flows per node.
These models do not account for non-independence between these and other flow properties.
We discuss this limitation and its implications when we analyze our experimental results.

@@

The striking results of applying nonnegative matrix factorization to $\mat{P}$ are shown in \Figure{basic-behaviors}. Twelve basic behaviors suffice to explain all the traffic in a ``toy''  sample of 4138 flows. The pairs of packet size and inter-packet interval distributions are almost immediately recognizable as resembling intuitive modes of network behavior. The first basic behavior, for example has all packets near the \caps{MTU}, while intervals are smoothly distributed around a fraction of a millisecond. This is what we would expect from downloading large files; and indeed, the breakdown of packets by port number bears out this interpretation---almost all of the traffic is \caps{HTTP} server traffic, with small slivers of \caps{IMAP} and \caps{SSH} server traffic. The second basic behavior is equally intuitive: tiny packets with frequency distributed around 0.1 seconds---this is all \caps{SSH} traffic, as indicated by the associate pie chart.
The other behaviors are listed in \Table{basic-behaviors} with interpretations.
% This is precisely the behavior we expect from interacting with a remote terminal via keystrokes and screen updates.

\input{tables/basic_behaviors}

@@

@@

@@

\bibliographystyle{plain}
\bibliography{references}
